# Merge three cycles

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
#library(dplyr)
library(readr)
library(DataExplorer)
library(tableone)
library(Publish)
library(jtools)
require(mice)
```

## Analytic dataset

### Load 2013-18 datasets

```{r, cache=TRUE}
load("data/analytic13recoded.RData")
load("data/analytic15recoded.RData")
load("data/analytic17recoded.RData")
```

### Merge 2013-18 datasets

```{r, cache=TRUE}
# adults aged 20 years or more
data.merged0 <- rbind(analytic13, analytic15, analytic17)
dim(data.merged0)
data.merged <- droplevels(data.merged0)
```

### Check missingness

```{r, cache=TRUE}
plot_missing(data.merged)
# profile_missing(data.merged)
dim(data.merged)
```

::: column-margin
The data contants variables with some missing information.
:::

```{r, cache=TRUE}
data.complete <- na.omit(data.merged)
dim(data.complete)
```

::: column-margin
-   Only complete cases retained, and survey features/weights were ignored for simplicity.
-   In a realistic analysis, we would consider the missingness pattern before deleting or imputing such information.
:::

## Summary statistics

```{r, cache=TRUE, echo=FALSE}
tab1 <- table1::table1(x = ~ age.cat + sex + education + race + 
                         marital + income + born + year + 
                         diabetes.family.history + smoking + diet.healthy + 
                         physical.activity + medical.access + sleep + 
                         systolicBP + diastolicBP + uric.acid + 
                         protein.total + bilirubin.total + phosphorus + 
                         sodium + potassium + globulin + calcium.total + 
                         high.cholesterol| obese, data = data.complete)
tab1
```

::: column-margin
-   Investigator specified covariates stratified by the exposure (obesity)
-   This Table includes information about participants with and without ICD-10-CM proxy information. Therefore, the sample is is larger than the original analysis.
:::

## Proxy data from ICD10 codes

```{r, cache=TRUE}
dat.proxy.long <- rbind(rx2013, rx2015, rx2017) 
dat.proxy.long$icd10 <- NULL
# Rename 3 digits ICD-10 codes as icd10
colnames(dat.proxy.long)[names(dat.proxy.long)=="icd10.new"] <- "icd10"
```

::: column-margin
We combine all of the ICD-10-CM information form all 3 cycles.
:::

## Save dataset for later use

```{r, cache=TRUE}
save(data.merged, 
     data.complete, 
     dat.proxy.long, 
     file = "data/analytic3cycles.RData")
```
