# Pure ML

::: callout-tip
We show an example using LASSO aproach
:::

# Pure ML approach (LASSO)

Start with all recurrence variables 


Say, 100 proxies (associated with outcome) were selected by LASSO approach (*ML-hdPS*)

```{r, cache=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
require(autoCovariateSelection)
require(dplyr)
require(cobalt)
library(ggplot2)
require(WeightIt)
library(lmtest)
library(knitr)
library(kableExtra)
load(file = "data/dfx7.RData")
```

## Choose variables associated with outcome

```{r, cache=TRUE, message=FALSE, warning=FALSE}
proxy.dim <- out2
dim(proxy.dim) # id + k
head(proxy.dim)[,1:3]
proxy.dim$id <- proxy.dim$idx
proxy.dim$idx <- NULL
fullcovproxy.data <- merge(data.complete[,c("id",
                                    outcome, 
                                    exposure, 
                                    investigator.specified.covariates)], 
                       proxy.dim, by = "id")
dim(fullcovproxy.data)
fullcovproxy.data$outcome <- as.numeric(I(fullcovproxy.data$diabetes=='Yes'))
fullcovproxy.data$exposure <- as.numeric(I(fullcovproxy.data$obese=='Yes'))
```


```{r, cache=TRUE, message=FALSE, warning=FALSE}
proxy.list <- names(out2[-1])
# out3$autoselected_covariate_df[,-1] for hybrid 
# out2 is from step2$recurrence_data
covarsTfull <- c(investigator.specified.covariates, proxy.list)
Y.form <- as.formula(paste0(c("outcome~ exposure", 
                              covarsTfull), collapse = "+") )
covar.mat <- model.matrix(Y.form, data = fullcovproxy.data)[,-1]
lasso.fit<-glmnet::cv.glmnet(y = fullcovproxy.data$outcome, 
                             x = covar.mat, 
                             type.measure='mse',
                             family="binomial",
                             alpha = 1, 
                             nfolds = 5)
coef.fit<-coef(lasso.fit,s='lambda.min',exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.ml <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.ml)
```

::: column-margin
-   From all proxies, we try to identify proxies that are empirically associated with the outcome based on a multivariate LASSO (outcome with all proxies in **one model**).
-   Note that LASSO model is choosing variables based on <mark>association with the `outcome`</mark> conditional on the 'exposure\`.
-   <mark>Variable selection</mark> is only happening for proxy variables.
-   Investigator specified variables are not being subject to variable selection.
:::

## Build model formula based on selected variables

```{r, cache=TRUE}
covform <- paste0(investigator.specified.covariates, collapse = "+")
proxyform <- paste0(proxy.list.sel.ml, collapse = "+")
rhsformula <- paste0(c(covform, proxyform), collapse = "+")
ps.formula <- as.formula(paste0("exposure", "~", rhsformula))
```

::: column-margin
Build propensity score model based on selected variables based on LASSO.
:::

## Fit the PS model

```{r, cache=TRUE}
require(WeightIt)
W.out <- weightit(ps.formula, 
                    data = fullcovproxy.data, 
                    estimand = "ATE",
                    method = "ps")
```

::: column-margin
Propensity score model fit to be able to calculate the inverse probability weights.
:::

## Obtain log-OR from unadjusted outcome model

```{r, cache=TRUE, warning=FALSE, message=FALSE}
out.formula <- as.formula(paste0("outcome", "~", "exposure"))
fit <- glm(out.formula,
            data = fullcovproxy.data,
            weights = W.out$weights,
            family= binomial(link = "logit"))
fit.summary <- summary(fit)$coef["exposure",
                                 c("Estimate", 
                                   "Std. Error", 
                                   "Pr(>|z|)")]
fit.ci <- confint(fit, level = 0.95)["exposure", ]
fit.summary_with_ci <- c(fit.summary, fit.ci)
knitr::kable(t(round(fit.summary_with_ci,2))) 
```

::: column-margin
Summary of results (log-OR).
:::

```{r, cache=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
save(dfx, patientIds, basetable, df, df2, step1, out1, step2, out2, out3,
     fullcovproxy.data, exposure, outcome, investigator.specified.covariates,
     proxyform, W.out, proxy.list, proxy.list.sel,fit.summary_with_ci,covform,
     file = "data/fullcovproxy.RData")
```
